### 2024-04-20 MSRI-16-1989-79.Rmd
gitignore

::: {.rmenv #chunk-1.2 title="1.2" latex="{1.2}"}
If we only consider smooth and projective varieties over a field $k$, and we assume the "standard" conjectures on algebraic cycles, then Grothendieck has shown how to define the category of motives generated by the $\HH_\mot^i(X)$ (cf. [@Kl; @Man]);
it is a semi-simple abelian category.

If we do not restrict ourselves to the category generated by the $\HH_\mot^i(X)$ for $X$ smooth and projective over $k$, then we no longer have even a conjectural definition of what the category of motives over $k$ should be.
However, the philosophy of motives is not made any less useful by this fact:
it organises known facts, poses questions, and suggests precise conjectures.
:::

::: {.rmenv #chunk-1.3 title="1.3" latex="{1.3}"}
\oldpage{92}In each of the known theories, the $\HH^i(X)$ are endowed with an increasing filtration $W$, known as the _weight filtration_ [@D5], as well as comparison isomorphisms such that (1.1.1) and (1.1.2) are compatible with $W$.
Furthermore, every natural map is strictly compatible with $W$.
From this, we get a new requirement for the category of motives:
every motive is endowed with a weight filtration $W$, compatible with the tensor product, and strictly compatible with every morphism $f\colon M\to N$, i.e.
$$
  f(M)\cap W_i(N)
  = f(W_i(M)).
$$

We say that a motive $M$ is _pure of weight $i$_ if $W_i(M)=M$ and $W_{i-1}(M)=0$.
For $X$ smooth and projective, $\HH_\mot^i(X)$ is pure of weight $i$.
We want for the $\otimes$-category generated by the $\HH_\mot^i(X)$, for $X$ smooth and projective over $k$, to be the sum of pure motives.
In terms of pure motives, the properties of $W$ can be written as follows:
every motive is the iterated extension of pure motives, and, for $M$ and $N$ pure of weights $m$ and $n$ (respectively),

a. $M\otimes N$ is pure of weight $m+n$;
b. for $m\neq n$, $\Hom(M,N)=0$; and
c. for $m\leqslant n$, $\Ext^1(M,N)=0$.

Often, pure motives (or direct sums of pure motives) are simply called _motives_, and their category admits the conjectural description [@Kl; @Man];
the more general motives, considered here, are then called _mixed motives_
:::

@@

_Terminology:_
all the vector spaces are assumed to be of finite dimension;
\oldpage{93}"_almost every_ prime number" means "all, except for a finite number".

>>>>

@@

_Terminology:_
all the vector spaces are assumed to be of finite dimension;
\oldpage{93}
"_almost every_ prime number" means "all, except for a finite number".
